---
title: "My M&M OCD"
author: "Yoni"
date: "`r format(Sys.time(), '%d %m, %Y')`"
output: pdf_document
---

# Intro

The goal of this simulation is to test the statistics of M&M and other stacks even Chocolate lentils by color,
I wanted to know, if I eat m&m package 2 by 2, separated by color, what is the chance of my finishing the package without mixing any color in one bite.

In addition, here are some BI incite that needed to be checked:

1. What is the probability of M&M packages packaged fairly?
2. What is the probability of M&M packages packaged without one color?
3. How does the size of the package or number of colors affect this probability?

###

The method is based of simulation of some M&M bags, according to the most common sizes.
Each time we sample x lentils, name them by colors (represented as factorial numbers), and see the results for many packages as a statistic data.

### Parameters

```{r packages, error=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(MASS) #calculate distribution parameters
library(scales)
```
The basic parameters (will be changed later):

```{r parameters, error=FALSE, warning=FALSE}
#parameters
nn<- 500            #numbers of bags per sample
n_color<- 6         #unique colors of M&M
gram<- 0.91         #weight of one M&M
bag_g<- 250         #common weight of M&M package
n_unit<- bag_g/gram #M&M per package
av_per_color= n_unit/n_color
paste0("The avarage number of lentils per color is ", round(av_per_color,2))
```

# Creating of Sample

## General Sample

create_bag is a function to create one snack package as matrix.

sample_MnM is a function to create n bags from the create_bag function.

```{r sample one bag, error=FALSE, warning=FALSE, echo=F}
#create bag using sample vector
create_bag<- function(x,tx)
  {
  x<- floor(x)+sample(0:1,1)
  res<- matrix(nrow = 1, ncol = tx)
  colnames(res)<- 1:tx
  S_res<- sample(tx,x,replace = T)
  while(sum(table(S_res)>x))
  {S_res<- S_res[-1]}
  for (i in 1:tx) {
    res[,i] <- sum(S_res==i)
    }
  res}

sample_MnM<-function(nn,x_units,t_colors)
  {
  res<- matrix(nrow = nn, ncol = t_colors)
  for (i in 1:nn)
    {res[i,]<- create_bag(x= x_units,tx=t_colors)}
  res}

```
```{r sample example, error=FALSE, warning=FALSE, echo=FALSE}
#example
print("One bag:")
create_bag(100,6)
print("3 bags:")
sample_MnM(3,10,6) %>% 
   `colnames<-`(c(1:6)) %>% 
   `rownames<-`(c("Bag_1","Bag_2","Bag_3"))
```

## Preview Graph

Now will be creating nn bugs of M&M

```{r creating the bag, warning=FALSE, error=FALSE}
MnM_sample<- sample_MnM(nn,n_unit,n_color)

MnM_sample<-
  MnM_sample %>% as_data_frame() %>% 
  mutate(even_count= rowSums(across(everything()  , ~ .x %% 2 == 0)), #how many evens colors there are
         even_evens= rowSums(across(c(1:6), ~ .x %% 2 == 1)) %% 2 ==0, #are the uneven colors even
         Variance=    apply(across(c(1:6)), 1, var), #var of candy per color/ type
         low_col=    rowSums(across(c(1:6), ~ .x <= 0.6*av_per_color)),
         min=        apply(across(c(1:6)),1,FUN = min)) # lowest value in color

MnM_sample %>% head(6)
```

ggplot the MnM sample sample

```{r ggplot 1 sample, warning=FALSE, error=FALSE}
MnM_sample$even_count %>% 
table()

MnM_sample %>% 
  mutate(is.even = even_count %%2 ==0) %>% 
  ggplot(aes(x= even_count, fill= is.even))+
  geom_bar()

MnM_sample %>%
  mutate(is.even = V1 %%2 ==0) %>% 
  ggplot(aes(x= V1, fill= is.even, alpha= 0.7))+
  geom_bar()

#summary of all colors
rbind(
MnM_sample$V1 %>% summary(),
MnM_sample$V2 %>% summary(),
MnM_sample$V3 %>% summary(),
MnM_sample$V4 %>% summary(),
MnM_sample$V5 %>% summary(),
MnM_sample$V6 %>% summary()
) %>% as.data.frame() %>% cbind(sapply(MnM_sample[,1:6],var,na.rm=1)) %>% 
  rename("Var" ="sapply(MnM_sample[, 1:6], var, na.rm = 1)")
```

## Test Expected Value
to see is the mu of the lentils per color are fair, we will test it per column with t.test.

```{r t test mu}
#test mu is av_per_color
check_mean_hypothesis <- function(data, column_name, X) {
  test_result <- t.test(data[[column_name]], mu = X)
  return(test_result$p.value)
}

columns_to_test <- colnames(MnM_sample[,1:6])
test_results <- sapply(columns_to_test, function(col) {
  check_mean_hypothesis(MnM_sample, col, av_per_color)
})

t(t(test_results))
```

```{r P.value test}
check_mean_hypothesis <- function(data, column_name, X) {
  t.test(data[[column_name]], mu = X)$p.value}

check_two_sample_t_test <- function(data, col1, col2) {
  t.test(data[[col1]], data[[col2]])$p.value}

columns_to_test <- colnames(MnM_sample[,1:6])
num_cols <- length(columns_to_test)
p_value_matrix <- matrix(NA, nrow = num_cols, ncol = num_cols, dimnames = list(columns_to_test, columns_to_test))
for (i in 1:num_cols) {
  for (j in 1:num_cols) {
    if (i == j) { #diagonal test of mu
      p_value_matrix[i, j] <- check_mean_hypothesis(MnM_sample, columns_to_test[i], av_per_color)
    } else { # 2 sample t.test
      p_value_matrix[i, j] <- check_two_sample_t_test(MnM_sample, columns_to_test[i], columns_to_test[j])
    }
  }
}

round(as.data.frame(p_value_matrix),4)
```

```{r plot 2 colors, warning=FALSE, error=FALSE}
MnM_sample %>%
    pivot_longer(cols = 1:6, names_to = "Variable", values_to = "pieces") %>%
    ggplot(aes(fill=Variable ,x= pieces, alpha= 0.4))+
  geom_histogram(position = "identity")
```

we can see that the variance distribution is Gamma like with shape and rate as seen below

```{r, warning=FALSE, error=FALSE}
gamma_params <- fitdistr(MnM_sample$Variance, "gamma")$estimate
paste0("the parameters of the gamma shaped var is shape ", round(gamma_params[1],3)," and rate ", round(gamma_params[2],3) )
shape_est <- gamma_params["shape"]
scale_est <- 1 / gamma_params["rate"] #using Scale= 1/Rate

MnM_sample %>% #plot density
  ggplot(aes(x = Variance, fill = "orange")) +
  geom_density(fill= "lightblue4", alpha= 0.8) +
  geom_vline(xintercept = mean(MnM_sample$Variance), color = "blue", size=1) +
  geom_vline(xintercept = scale_est*shape_est, color = "red", linetype = "dashed", size=1) +
  stat_function(fun = dgamma, args = list(shape = shape_est, scale = scale_est), 
                color = "red", alpha= 0.7) +
  labs(title = "Density Plot with Gamma Distribution",
       x = "Variance",
       y = "") +
  scale_y_continuous(label=scales::label_percent(.1)) + 
  theme_minimal()
```

#check too low color (under 10%) and sample by n number

#use statistics to sample better low chance cases

# optimizing best package

### n*m types of snacks

We will create a function that create sample for each number of colors and package size we want, and then calculate some interesting parameters

```{r function multy-sample, warning=FALSE, error=FALSE}
mega_snack<- function(nn,n_unit,n_color)
  {
  m_sample<- length(n_unit)*length(n_color)
  nul_mat= matrix(,nrow = m_sample, ncol = 5)
  res<- cbind(rep(n_unit,length(n_color)),sort(rep(n_color,length(n_unit))),
                                               nul_mat)
  for (i in 1:(dim(res)[1]))
  {
    #print(c(res[i,1],res[i,2]))
    low_color<- 0.666*res[i,1]/(res[i,2])
    small_sample<- sample_MnM(nn,res[i,1],res[i,2])
    small_sample<-
      small_sample %>% as_data_frame() %>% 
      mutate(even_count= rowSums(across(everything()  , ~ .x %% 2 == 0))/n_color, #how many evens colors there are
             even_evens= (rowSums(across(c(1:n_color)  , ~ .x %% 2 == 1)) %% 2 ==0)/n_color, #are the uneven colors even
             var_col=    apply(across(c(1:n_color)), 1, var), #var of candy per color/ type
             all_even=   rowSums(across(c(1:n_color)  , ~ .x %% 2 == 0))== n_color,
             low_col=   rowSums(across(c(1:n_color), ~ .x <= low_color ))>=1)
    res[i,3]<- mean(small_sample$even_count)
    res[i,4]<- mean(small_sample$even_evens)
    res[i,5]<- mean(small_sample$var_col)
    res[i,6]<- mean(small_sample$all_even)
    res[i,7]<- mean(small_sample$low_col)
      }
  colnames(res)<- c("n_unit", "n_color", "even_count", "even_evens", "var_col","all_even","low_color")
  res
  }
```


```{r baking multy sample, warning=FALSE, error=FALSE}
color_op<- 2:7
grams_op<- c(25,45,150,250,330,500,750,1000)
n_unit_op<- grams_op/gram

mega_snack_1<-
  mega_snack(50,n_unit_op,color_op) %>% as.data.frame() %>% 
  mutate(n_unit= round(n_unit,1))

mega_snack_1 %>% head(6)
```

```{r hitmap, warning=FALSE, error=FALSE}
mega_snack_1 %>%
  ggplot(aes(x = factor((round( n_unit,1) )), y = factor(n_color ), fill = even_count )) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red4")+
  labs(title = "Heatmap of mega_snack even_count",
       x = "lentils",
       y = "No. of colors",
       fill = "even_count") +
  theme_minimal()
```
```{r hitmap_2, warning=FALSE, error=FALSE}
mega_snack_1 %>%
  ggplot(aes(x = factor((round( n_unit,1) )), y = factor(n_color ), fill = even_evens )) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "red")+
  labs(title = "Heatmap of mega_snack even_evens",
       x = "lentils",
       y = "No. of colors",
       fill = "even_evens") +
  theme_minimal()
```

now let us see the probability of all even, and whether there is pattern.
```{r ggplot all even, warning=FALSE, error=FALSE}
mega_snack_1 %>%
  ggplot(aes(x = (round( n_unit,1) ), y = n_color , color = all_even,size = all_even )) +
  geom_point() +
  scale_color_gradient(low = "orange3", high = "green4")
  labs(title = "Heatmap of mega_snack all evens",
       x = "lentils",
       y = "No. of colors",
       color = "all_even") +
  theme_minimal()
```

```{r ggplot missing color, warning=FALSE, error=FALSE}
mega_snack_1 %>%
  ggplot(aes(x = factor(round(n_unit, 1)), y = n_color, fill = low_color))+
  geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.2f%%", low_color * 100)), color = "black", size = 4)+
  scale_fill_gradient(low = "pink2", high = "red4",labels = scales::percent)+
  labs(title = "Chance of Less Than 66% Color in Package",
       x = "lentils",
       y = "No. of colors",
       fill = "low color")+
  theme_minimal()
```

As we can see, only the small package (less than 50 lentils) have high probability of at least one color to appear severely lower.

Therefore, splitting package by color on the big ones should be relatively even.
