---
title: "My M&M OCD"
author: "Yoni"
date: "`r format(Sys.time(), '%d %m, %Y')`"
output: pdf_document
---

# Intro

### Objective of Simulation

The goal of this simulation is to test the statistics of M&M and other stacks even Chocolate lentils by color.

I wanted to know what is the chance of my finishing the package of M&M without mixing any color in one bite, eating them 2 by 2

In addition, here are some BI incite that needed to be checked:

1. What is the probability of M&M packages packaged fairly?

2. What is the probability of M&M packages packaged without one color?

3. How does the size of the package or number of colors affect this probability?

### Method

I do not have an inner information of how does (M&M)[https://www.mms.com/en-us] make their delicious snacks nor we know how they make sure each package have fair amount of each color.
Therefore, the method I chose is based of simulation of some M&M bags, according to the most common sizes of packages.

Each time we sample x lentils(units of M&M), name them by colors (V1,V2...), and see the results for many packages as a statistic data.

my hypothesis is that the probability of perfect package (aka a package with all colors number been even) is very small, at least for a standard 6 colors pack.

### Parameters

Basic parameters:

```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(MASS) #calculate distribution parameters
library(scales)
```
```{r parameters, error=FALSE, warning=FALSE}
#parameters
n<- 800            #numbers of bags per sample
n_color<- 6         #unique colors of M&M
gram<- 0.91         #weight of one M&M
bag_g<- 250         #common weight of M&M package
n_unit<- bag_g/gram #M&M per packagenm,
av_per_color= n_unit/n_color
paste0("The avarage number of lentils per color is ", round(av_per_color,2))
```

# Creating of the Sample

## General Sample

In order to test the theoretical data, we need to simulate it using customize functions. here are there:

* Create_bag- function to create one snack package for chosen package size and number of colors.

* sample_MnM- function to create n bags from the Create_bag function.

```{r sampling functions, error=FALSE, warning=FALSE, echo=F}
Create_bag<- function(x,tx) #create bag using sample vector
  {
  x<- floor(x)+sample(0:1,1)
  res<- matrix(nrow = 1, ncol = tx)
  colnames(res)<- 1:tx
  S_res<- sample(tx,x,replace = T)
  while(sum(table(S_res)>x))
  {S_res<- S_res[-1]}
  for (i in 1:tx) {
    res[,i] <- sum(S_res==i)
    }
  res}

sample_MnM<-function(n,x_units,t_colors) #create n bag using sample vector
  {
  res<- matrix(nrow = n, ncol = t_colors)
  for (i in 1:n)
    {res[i,]<- Create_bag(x= x_units,tx=t_colors)}
  res}
```
```{r sample example, error=FALSE, warning=FALSE, echo=FALSE}
#example
print("One bag:")
Create_bag(100,6)
print("3 bags:")
sample_MnM(3,10,6) %>% 
   `colnames<-`(c(1:6)) %>% 
   `rownames<-`(c("Bag_1","Bag_2","Bag_3")) %>% 
   `colnames<-`(c("Red","Blue","Green", "Orange", "Yellow", "Brown"))
```

## Preview Graph

Now will be creating n bugs of M&M

columns:

1. V1:V6- the number of lentils per color

2. even_count- how many evens colors there are

3. even_evens- are the uneven colors even

4. Variance- variance of lentils per color

5. low_col- sum true if one color's count is lower than $\frac{2}{3}$ of expected value

6. min- the lowest color in each row

here are the first rows:

```{r creating the bag, warning=FALSE, error=FALSE, echo=FALSE}
MnM_sample<- sample_MnM(n,n_unit,n_color) %>% 
     `colnames<-`(c("Red","Blue","Green", "Orange", "Yellow", "Brown"))

MnM_sample<-
  MnM_sample %>% as_data_frame() %>% 
  mutate(even_count= rowSums(across(everything()  , ~ .x %% 2 == 0)), #how many evens colors there are
         even_evens= rowSums(across(c(1:6), ~ .x %% 2 == 1)) %% 2 ==0, #are the uneven colors even
         low_col=    rowSums(across(c(1:6), ~ .x <= 0.66*av_per_color)),
         Variance=    apply(across(c(1:6)), 1, var), #var of candy per color/ type
         min=        apply(across(c(1:6)),1,FUN = min), # lowest value in color
         all_even=   rowSums(across(c(1:n_color)  , ~ .x %% 2 == 0))== n_color)


MnM_sample %>% head(6)

```

plot the M&M sample sample

```{r ggplot 1 sample, warning=FALSE, error=FALSE, echo=F}
print("summary of all colors Distibution:")  #summary of all colors

rbind(
MnM_sample$Red %>% summary(),
MnM_sample$Blue %>% summary(),
MnM_sample$Green %>% summary(),
MnM_sample$Orange %>% summary(),
MnM_sample$Yellow %>% summary(),
MnM_sample$Brown %>% summary()
) %>% as.data.frame() %>% cbind(sapply(MnM_sample[,1:6],var,na.rm=1)) %>% 
  rename("Var" ="sapply(MnM_sample[, 1:6], var, na.rm = 1)")

MnM_sample %>%
  mutate(is.even = even_count %%2 ==0) %>% 
  ggplot(aes(x= even_count, fill= is.even))+
  geom_bar()+
    geom_text(
    stat = "count",
    aes(label = ..count..),vjust = -0.5
  )+labs(title="Distibution of Eveness of Colors")

MnM_sample %>%
  mutate(is.even = Red %%2 ==0) %>% 
  ggplot(aes(x= Red, fill= is.even))+
  scale_fill_manual(values = c("red", "red4"))+
  geom_bar()+labs(title="Distibution of the Red Color")+
  theme(panel.grid = element_line(color = "gray50"))
```
## Statisics Checking of the Simullation

### Test Expected Value

to see is the $\mu$ of the lentils per color are fair, we will test it per column with t.test for each color.

Here is the result, none of them bellow 5% P. value

```{r t test mu, echo=FALSE}
#test mu is av_per_color
check_mean_hypothesis <- function(data, column_name, X)
  {
  test_result <- t.test(data[[column_name]], mu = X)
  test_result$p.value
}

columns_to_test <- colnames(MnM_sample[,1:6])
test_results <- sapply(columns_to_test, function(col) {
  check_mean_hypothesis(MnM_sample, col, av_per_color)
})

percent(test_results)
```

Now we will do the same checking for 2 samples, to see whether there is correlation between each 2 colors distribution.

for each row i and column j,
1) if i==j, this it the check from before of the expected value to n_unit/n_color
2) if i!=j, this is two samples test of same expected value hypothesis

```{r P.value test, echo=FALSE}
check_mean_hypothesis <- function(data, column_name, X) {
  t.test(data[[column_name]], mu = X)$p.value}

check_two_sample_t_test <- function(data, col1, col2) {
  t.test(data[[col1]], data[[col2]])$p.value}

columns_to_test <- colnames(MnM_sample[,1:6])
num_cols <- length(columns_to_test)
p_value_matrix <- matrix(NA, nrow = num_cols, ncol = num_cols, dimnames = list(columns_to_test, columns_to_test))
for (i in 1:num_cols) {
  for (j in 1:num_cols) {
    if (i == j) { #diagonal test of mu
      p_value_matrix[i, j] <- check_mean_hypothesis(MnM_sample, columns_to_test[i], av_per_color)
    } else { # 2 sample t.test
      p_value_matrix[i, j] <- check_two_sample_t_test(MnM_sample, columns_to_test[i], columns_to_test[j])
    }
  }
}

round(as.data.frame(p_value_matrix),4) %>%
  mutate(color_1= rownames(p_value_matrix)) %>% 
  pivot_longer(cols = 1:6, names_to = "color_2", values_to = "t_test") %>% 
  mutate(reject_H0 = t_test<= 0.05) %>% 
  ggplot(aes(x= color_1, y= color_2))+
  geom_text(aes(label = t_test * 100,, color = reject_H0), size = 4)+
  labs(title="Colors Correlation Map")
```
Now here Is visualization of the actual data per color

```{r plot 2 colors, warning=FALSE, error=FALSE, echo=F}
MnM_sample %>%
    pivot_longer(cols = 1:6, names_to = "Variable", values_to = "pieces") %>%
    ggplot(aes(fill=Variable ,x= pieces))+
  geom_histogram(position = "identity")+
    scale_fill_manual(values = c("blue", "chocolate4","green4", "orange","red4", "yellow3"))+
  theme(panel.grid = element_line(color = "grey50",
                                  size = 0.75))+facet_wrap(~Variable)
```

### Variance Distribution Checking

We know that the distribution of variance is approximately Gamma distribution:

$$f(x) = \frac{1}{(\Gamma(\alpha)\theta^\alpha)} x^{\alpha-1}e^{-x/\theta}$$

We can see that the variance distribution is Gamma like with shape and rate as seen below

```{r, warning=FALSE, error=FALSE, echo=F}
gamma_params <- fitdistr(MnM_sample$Variance, "gamma")$estimate
paste0("The parameters of the gamma shaped variance is shape ", round(gamma_params[1],3)," and rate ", round(gamma_params[2],3) )
shape_est <- gamma_params["shape"]
scale_est <- 1 / gamma_params["rate"] #using Scale= 1/Rate

MnM_sample %>% #plot density
  ggplot(aes(x = Variance)) +
  geom_density(fill= "skyblue", alpha= 0.8) +
  geom_vline(xintercept = mean(MnM_sample$Variance), color = "blue", size=1) +
  geom_vline(xintercept = scale_est*shape_est, color = "red", linetype = "dashed", size=1) +
  stat_function(fun = dgamma, args = list(shape = shape_est, scale = scale_est), 
                color = "red", alpha= 0.7) +
  labs(title = "Density Plot with Gamma Distribution",
       x = "Variance",
       y = "") +
  scale_y_continuous(label=scales::label_percent(.1)) + 
  theme_minimal()
```
### Are All Even in the Sample?

```{r are all even, warning=FALSE, error=FALSE, echo=F}
MnM_sample %>%
  select_(1,2,3,4,5,6) %>%
  mutate(count_even = rowSums(across(1:n_color, ~ .x %% 2 == 0))) %>%
  count(count_even) %>%
  ggplot(aes(x = factor(count_even), y = n, fill = n)) +  # make x categorical
  geom_bar(stat = "identity", color= "grey20") +
  scale_fill_gradient(low = "skyblue", high = "green3") +
  labs(
    title = "Count of Even Colors per Row",
    x = "No. Even Rows",
    y = "Packages"
  )+
    geom_text(aes(label = n), color = "black", size = 4)+
  theme_minimal()
```

### n*m types of snacks

I will create a function that create sample for each number of colors and package size we want, and then calculate some interesting parameters

```{r function multy-sample, warning=FALSE, error=FALSE, echo=F}
mega_snack<- function(nn,n_unit,n_color)
  {
  m_sample<- length(n_unit)*length(n_color)
  nul_mat= matrix(nrow = m_sample, ncol = 6)
  res<- cbind(rep(n_unit,length(n_color)),
              sort(rep(n_color,length(n_unit))),
              nul_mat)
    colnames(res)<- c("n_unit", "n_color", "even_count", "even_evens",
                    "var_col","all_even","low_color","smallest_col")
    for (i in 1:(dim(res)[1]))
  {
    low_color<- 0.666*res[i,1]/(res[i,2])
    small_sample<- sample_MnM(nn,res[i,1],res[i,2])%>%
      as_data_frame() %>% 
      mutate(even_count= rowSums(across(everything()  , ~ .x %% 2 == 0))/res[i,2],
             #how many evens colors there are
             even_evens= (rowSums(across(c(1:n_color)  , ~ .x %% 2 == 1)) %% 2 ==0)/res[i,2],
             #are the uneven colors even
             var_col=     apply(across(c(1:n_color)), 1, var),
             #var of candy per color/ type
             smallest_col=apply(across(c(1:n_color)), 1, min),
             #lowest value in color
             #all_even=    rowSums(across(c(1:n_color), ~ .x %% 2 == 0)) >= res[i,2],
             all_even=    even_count >= 1,
             low_color=   rowSums(across(c(1:n_color), ~ .x <= low_color ))>=1
            )
    #return(small_sample)
    res[i,3:8]<- c(
      mean(small_sample$even_count),
      mean(small_sample$even_evens),
      mean(small_sample$var_col),
      mean(small_sample$all_even,  na.rm = T),
      mean(small_sample$low_color),
      min(small_sample$smallest_col))
      }

  res %>% as.data.frame()
}
```


```{r Data for multiple sample, warning=FALSE, error=FALSE}
color_op<- 2:8
grams_op<- c(25,45,150,250,330,500,750,1000)
n_unit_op<- grams_op/gram
nn=800
```

We will make the multiple sample. Here is some random rows:

```{r baking multiple sample, warning=FALSE, error=FALSE, echo= F}
mega_snack_1<-
  mega_snack(nn,n_unit_op,color_op) %>%
  mutate(n_unit= round(n_unit,1))
mega_snack_1 %>% sample_n(5)
```

## Deep Insight on the Data

here are some insights:

```{r hitmap, warning=FALSE, error=FALSE, echo=F}
mega_snack_1 %>%
  ggplot(aes(x = factor((round( n_unit,1) )), y = factor(n_color ), fill = even_count)) +
  geom_tile(color= "black") +
  scale_fill_gradient(low = "orange", high = "skyblue")+
  labs(title = "avarage even count percent by number of colors and units",
       x = "Lentils",
       y = "No. of colors",
       fill = "even_count") +
  guides(colour= "none")+
  theme_minimal()
```


```{r hitmap_2, warning=FALSE, error=FALSE, echo=F}
mega_snack_1 %>%
  ggplot(aes(x = factor((round( n_unit,1) )), y = factor(n_color ), fill = even_evens )) +
  geom_tile(color= "black") +
  scale_fill_gradient(high = "deepskyblue", low = "red")+
  labs(title = "does the Uneven Colors Even",
       x = "Lentils",
       y = "No. of colors",
       fill = "even evens") +
  geom_text(aes(label = round(even_evens,2)), color = "black", size = 4)+
  theme_minimal()
```

Here is probability of all even, and whether there is pattern.

```{r ggplot all even, warning=FALSE, error=FALSE, echo=F}
Heatmap_all_evens<-
  mega_snack_1 %>%
  ggplot(aes(x = as.factor(round( n_unit,1) ), y = n_color , color = all_even)) +
  geom_point(size = 8 ) +
  scale_color_gradient(low = "yellow", high = "darkolivegreen3")+
  labs(title = "Were all Colors Evens",
       x = "Lentils",
       y = "No. of colors",
       color = "all_even") +
        geom_text(aes(label = sprintf("%.1f%%", all_even * 100)), color = "black", size = 3)+
    guides(size= "none")+
  theme_bw()

ggsave(Heatmap_all_evens, 
       filename = "Heatmap_all_evens.png",
       height = 4, width = 5, units = "in")
  guides(colour= "none")

Heatmap_all_evens
```

```{r ggplot missing color, warning=FALSE, error=FALSE, echo=F}
mega_snack_1 %>%
  ggplot(aes(x = factor(round(n_unit, 1)), y = n_color, fill = low_color))+
  geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.2f%%", low_color * 100)), color = "black", size = 4)+
  scale_fill_gradient(low = "darkolivegreen3", high = "red3",labels = scales::percent)+
  labs(title = "Chance of Less Than 66% Color in Package",
       x = "Lentils",
       y = "No. of colors",
       fill = "low color")+
  theme_minimal()
```

```{r smallest color vs expected, warning=FALSE, error=FALSE, echo=F}
mega_snack_1 %>% 
  select_(1,2,8) %>% 
  mutate(smallest_to_mu= smallest_col*n_color/n_unit) %>%
   ggplot(aes(x = factor(round(n_unit, 1)), y = n_color, fill = smallest_to_mu))+
  geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.1f%%", smallest_to_mu * 100)), color = "black", size = 4)+
  scale_fill_gradient(low = "red3", high = "darkolivegreen3",labels = scales::percent)+
  labs(title = "Smallest Color as Percent of Expected Value",
       x = "Lentils",
       y = "No. of colors",
       fill = "% of Expected Value")+
  theme_bw()
```

As we can see, only the small package (less than 50 lentils) have high probability of at least one color to appear severely lower.

Therefore, splitting package by color on the big ones should be relatively even.

### using regression for correlation check

```{r regression, warning=FALSE, error=FALSE, echo=F}
mega_snack_open<- function(nn,n_unit,n_color)
  {
  m_sample<- nn*length(n_unit)*length(n_color)
  nul_mat= matrix(nrow = m_sample, ncol = 2)
  res<- cbind(rep(n_unit,nn*length(n_color)),
         sort(rep(n_color,nn*length(n_unit))),
              nul_mat)
  for (i in 1:(dim(res)[1]))
  {
    low_color<- 0.666*res[i,1]/(res[i,2])
    small_sample<- Create_bag(res[i,1],res[i,2])%>%
      as_data_frame() %>% 
      mutate(even_count= rowSums(across(everything()  , ~ .x %% 2 == 0))/res[i,2],
             all_even=   rowSums(across(c(1:n_color)  , ~ .x %% 2 == 0))== res[i,2],
            )
    res[i,3]<- mean(small_sample$even_count)
    res[i,4]<- mean(small_sample$all_even)
      }
  colnames(res)<- c("n_unit", "n_color", "even_count","all_even")
  res %>% as.data.frame()
  }

mega_snack_2<- mega_snack_open(15,n_unit_op,color_op) %>% 
  mutate(color_No2= n_color %%2 == 0) #check if the number of colors is even
lm_color<- lm(mega_snack_2$even_count~ mega_snack_2$n_color +mega_snack_2$n_unit)
lm_color %>% summary()


lm_color_all<- lm(mega_snack_2$all_even~ mega_snack_2$n_color +mega_snack_2$n_unit+ mega_snack_2$color_No2)
lm_color_all %>% summary()
```

# Conclusions

### Data Structure

The simulation created a random samples of snack packs, which was proven to be statistically random. We created with "sample_MnM" one sample with specific size and numbers of colors, and then "mega_snack" that create costume samples and check the relevand indicators fot this project.

we I out that:

* the smallest package can barely have one color or even to not have one.

* The more colors there are, the less chance there is for the uneven numbers in colors to be even

* For any medium or bigger package, the probability of one color to be unfairly small (less than $\frac{2}{3}$ than expected) is nearly 0%

### Main Q: Eating M&M by Two

Although there is no clear pattern to the right M&M package for all the colors to have even count,
maybe different approach can find a clear reason for more or less couples of M&M.

The general probability of all colors to be even in 6 colored pack is 1.5% for small 50g package 2.1% for big 1000g package, and overall 1.5%, which is more than I expected.

For 5 colored pack like Skittles the average is about 2.9%

For 2 colored pack the average is 24.5%, so for most 2 colored marshmallow bag the method of eating by 2 can be relatively available.

### Summery

To sum it up, for each medium pack the probability of all even colors is 1.4%, or 1 in a 73 packs of 250g. So I might need to change my snack preference to marshmallow if I want to keep this method!
