---
title: "Statistical Analysis of Random Distribution in M&M Packages"
author: "Yoni"

date: "`r format(Sys.time(), '%d %m, %Y')`"
output: github_document
always_allow_html: true
---

# Intro

### Objective of Simulation

This simulation explores the fairness and randomness in the color distribution of M&M-style chocolate candies. Specifically, it investigates:

1.  How often are M&M packs evenly distributed across all colors?

2.  What’s the probability of a pack missing at least one color?

3.  How do the number of candies and available colors impact that probability?

while seeing some chocolate packages, I wondered:

      **Can I finish a pack of M&M eating two at a time, without ever mixing colors in a bite?**

This playful question leads to a deeper statistical exploration of random sampling, distribution fairness, and packaging quality in candy production.


### Methodology

Since production data from [M&M](https://www.mms.com/en-us) isn’t publicly available, I simulate packages based on common retail sizes and the standard 6-color set. Each simulation randomly draws a sample of candies, assigning each a color. We repeat this process hundreds of times to analyze statistical properties across "virtual" packages.

Through this approach, I estimate the probability of:

1.  Getting a perfect pack (i.e., equal counts of each color).

2.  Receiving a pack missing at least one color.

3.  Seeing how these probabilities change with pack size or number of available colors.

My hypothesis is that perfectly balanced packs are extremely rare, especially when six colors or more are involved. Larger packs may contain all colors more consistently but still tend to be uneven in distribution.

### Parameters

Basic parameters:

We define a "pack" as a vector of integers representing the count of each color.  
Each simulation uses random sampling with replacement to mimic real-world packaging.  
Key variables:

- `n-color`: Number of distinct colors

- `n-unit`: Total candies in the pack

```{r parameters setup}
#parameters
n= 1000              #numbers of bags per sample
n_color= 6          #unique colors of M&M
gram= 0.91          #weight of one M&M
bag_g= 250          #common weight of M&M package
n_unit= bag_g/gram  #M&M per package
```
```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(MASS) #calculate distribution parameters
library(scales)
library(patchwork) #combine 2 plots
library(stargazer) #regression presentation
```
```{r mu of units, error=FALSE, warning=FALSE, echo=FALSE}
av_per_color= n_unit/n_color
paste0("The avarage number of lentils per color is ", round(av_per_color,1))
```

# Creating the Sample

## General Sample

In order to test the theoretical data, I need to simulate it using customize functions. here are there:

* `Create_bag`- function to create one snack package for chosen package size and number of colors.

* `sample_MnM`- function to create n bags from the Create_bag function.

key parameters for `sample_MnM`

- `n`: Number of packages in the sample

- `x_units`: Total candies in each package

- `n_colors`: Number of distinct colors in each package

```{r sampling functions, error=FALSE, warning=FALSE, echo=F}
Create_bag<- function(bag,n_colors) #create bag using sample vector
  {
  x<- floor(bag)+sample(0:1,1)
  res<- matrix(nrow = 1, ncol = n_colors)
  colnames(res)<- 1:n_colors
  S_res<- sample(n_colors,x,replace = T)
  while(sum(table(S_res)>x))
  {S_res<- S_res[-1]}
  for (i in 1:n_colors) {
    res[,i] <- sum(S_res==i)
    }
  res}

sample_MnM<-function(n,x_units,n_colors) #create n bag using sample vector
  {
  res<- matrix(nrow = n, ncol = n_colors)
  for (i in 1:n)
    {res[i,]<- Create_bag(x_units,n_colors)}
  res}
```
```{r sample example, error=FALSE, warning=FALSE, echo=FALSE}
#example
print("One bag of 100:")
Create_bag(100,6)
print("3 bags of 100:")
sample_MnM(3,100,6) %>% 
   `colnames<-`(c(1:6)) %>% 
   `rownames<-`(c("Bag_1","Bag_2","Bag_3")) %>% 
   `colnames<-`(c("Red","Blue","Green", "Orange", "Yellow", "Brown"))
```

## Preview Graph

Now will be creating n bugs of M&M

columns:

1. `V1:V6`- the number of lentils per color

2. `even_count`- how many evens colors there are

3. `even_evens`- are the uneven colors even

4. `low_col`- sum true if one color's count is lower than $\frac{2}{3}$ of expected value

5. `Variance`- variance of lentils per color

6. `min`- the lowest color in each row

7. `all_even`- are all colors even

here are the first rows:

```{r creating the bag, warning=FALSE, error=FALSE, echo=FALSE}
MnM_sample<- sample_MnM(n,n_unit,n_color) %>% 
     `colnames<-`(c("Red","Blue","Green", "Orange", "Yellow", "Brown"))

MnM_sample<-
  MnM_sample %>% as_data_frame() %>% 
  mutate(even_count= rowSums(across(c(1:6), ~ .x %% 2 == 0)), #count evens colors
         even_evens= rowSums(across(c(1:6), ~ .x %% 2 == 1)) %% 2 ==0, #are the uneven colors even
         low_col=    rowSums(across(c(1:6), ~ .x <= 0.66*av_per_color)),
         Variance=    apply(across(c(1:6)), 1, var) %>% round(.,2), #var of candy per color/ type
         min=         apply(across(c(1:6)),1,FUN = min), # lowest value in color
         all_even=   even_count== 6
         )

knitr::kable(MnM_sample %>% head(4), caption = "M&M sample random rows")
```

I summarized the sample by color bellow

```{r ggplot colors summary df, warning=FALSE, error=FALSE, echo=F}
colors_summary_df<-rbind(
  MnM_sample$Red %>% summary(),
  MnM_sample$Blue %>% summary(),
  MnM_sample$Green %>% summary(),
  MnM_sample$Orange %>% summary(),
  MnM_sample$Yellow %>% summary(),
  MnM_sample$Brown %>% summary()) %>%
  as.data.frame() %>% cbind(sapply(MnM_sample[,1:6],var,na.rm=1) ) %>% 
  rename("Var" ="sapply(MnM_sample[, 1:6], var, na.rm = 1)") %>% 
  mutate(Var= round(Var,2))

knitr::kable(colors_summary_df, caption = "summary of all colors Distibution")
```

Here we can see the distribution of all colors to be even and of one example color (red)

```{r ggplot all colors and red, warning=FALSE, error=FALSE, echo=F}

red_1<- MnM_sample %>%
  mutate(is.even = even_count %%2 ==0) %>% 
  ggplot(aes(x= even_count, fill= is.even))+
  scale_fill_manual(values = c("darkorange3", "chocolate4"))+
  geom_bar()+
    geom_text(stat = "count",
    aes(label = ..count..),vjust = -0.5)+
  labs(title="Distibution of Eveness of Colors",
       x= "count of evens",
       y= "count",
       fill= "is even"
       )
  
red_2<- MnM_sample %>%
  mutate(is.even = Red %%2 ==0) %>% 
  ggplot(aes(x= Red, fill= is.even))+
  scale_fill_manual(values = c("red", "red4"))+
  geom_bar()+
  labs(title="Distibution of the Red Color",
       x= "No. of red",
       y= "count",
       fill= "is even")+
  theme(panel.grid = element_line(color = "gray50"))

red_1+red_2
detach("package:patchwork", unload=TRUE)
```

## Statisics Checking of the Simullation

### Test Expected Value

In order to see is the $\mu$ of the lentils per color are fair, I will test it per column with t.test for each color.

Here is the result, none of them bellow 5% P. value

p.value of $H_0: \mu= \frac{n-unit}{n-color}$

```{r t test mu, echo=FALSE}
#test mu is av_per_color
check_mean_hypothesis <- function(data, column_name, X)
  {
  test_result <- t.test(data[[column_name]], mu = X)
  test_result$p.value}

columns_to_test <- colnames(MnM_sample[,1:6])
test_results <- sapply(columns_to_test, function(col) {
  check_mean_hypothesis(MnM_sample, col, av_per_color)})

percent(test_results)
```

Now I will do the same checking for 2 samples, to see whether there is correlation between each 2 colors distribution.

for each row i and column j

1) if i==j, this it the check from before of the expected value to n_unit/n_color

2) if i!=j, this is two samples test of same expected value hypothesis

```{r P.value test, echo=FALSE}
check_mean_hypothesis <- function(data, column_name, X) {
  t.test(data[[column_name]], mu = X)$p.value}

check_two_sample_t_test <- function(data, col1, col2) {
  t.test(data[[col1]], data[[col2]])$p.value}

columns_to_test <- colnames(MnM_sample[,1:6])
num_cols <- length(columns_to_test)
p_value_matrix <- matrix(NA, nrow = num_cols, ncol = num_cols, dimnames = list(columns_to_test, columns_to_test))
for (i in 1:num_cols) {
  for (j in 1:num_cols) {
    if (i == j) { #diagonal test of mu
      p_value_matrix[i, j] <- check_mean_hypothesis(MnM_sample, columns_to_test[i], av_per_color)
    } else { # 2 sample t.test
      p_value_matrix[i, j] <- check_two_sample_t_test(MnM_sample, columns_to_test[i], columns_to_test[j])
    }
  }
}

round(as.data.frame(p_value_matrix),3) %>%
  mutate(color_1= rownames(p_value_matrix)) %>% 
  pivot_longer(cols = 1:6, names_to = "color_2", values_to = "t_test") %>% 
  mutate(Reject_H0 = t_test<= 0.05) %>% 
  ggplot(aes(x= color_1, y= color_2))+
  geom_text(aes(label = sprintf("%.1f%%", t_test * 100), color = Reject_H0), size = 4)+
  labs(title="Colors Correlation Map", color = "Reject Ho",
       x = "Color 1",
       y = "Color 2")+
  theme_classic()+
    theme(
  panel.grid.major.y = element_line(color = "grey80"),
  panel.grid.major.x = element_line(color = "grey80")
    )
```
Now here Is visualization of the actual data per color

```{r plot 2 colors, warning=FALSE, error=FALSE, echo=F, message= F}
colored_MnM<- MnM_sample %>%
    pivot_longer(cols = 1:6, names_to = "Colour", values_to = "Lentils") %>%
    ggplot(aes(fill=Colour ,x= Lentils))+
  geom_histogram(position = "identity", alpha= 0.8)+
    scale_fill_manual(values = c("blue", "chocolate4","green4", "orange3","red4", "yellow2"))+
  theme(panel.grid = element_line(color = "grey70",
                                  size = 0.75))+facet_wrap(~Colour)

ggsave(colored_MnM, 
       filename = "colored_M&M.png",
       height = 4, width = 5, units = "in")

colored_MnM
```

### Variance Distribution Checking

I know that the distribution of variance is approximately Gamma distribution:

$$f(x) = \frac{1}{\Gamma(\alpha)\theta^\alpha} x^{\alpha-1}e^{-x/\theta}$$

I can see that the variance distribution is Gamma like with shape and rate as seen below

```{r, warning=FALSE, error=FALSE, echo=F}
gamma_params <- fitdistr(MnM_sample$Variance, "gamma")$estimate
shape_est <- gamma_params["shape"]
scale_est <- 1 / gamma_params["rate"] #using Scale= 1/Rate
paste0("The parameters of the gamma shaped variance is shape ", round(shape_est,3)," and scale ", round(scale_est,3) )

MnM_sample %>% #plot density
  ggplot(aes(x = Variance)) +
  geom_density(fill= "skyblue3", alpha= 0.8) +
  geom_vline(xintercept = mean(MnM_sample$Variance), color = "blue", size=1) +
  geom_vline(xintercept = scale_est*shape_est, color = "red", linetype = "dashed", size=1) +
  stat_function(fun = dgamma, args = list(shape = shape_est, scale = scale_est), 
                color = "red3", alpha= 0.7) +
  labs(title = "Color Variance vs Gamma Distribution",
       subtitle = "comparing Variance of lentils colors to equivalent Gamma distibution",
       x = "Variance",
       y = "") +
  scale_y_continuous(label=scales::label_percent(.1)) + 
  theme_minimal()
```

### Are All Even in the Sample?

```{r are all even, warning=FALSE, error=FALSE, echo=F}
MnM_sample %>%
  select_(1,2,3,4,5,6) %>%
  mutate(count_even = rowSums(across(1:6, ~ .x %% 2 == 0))) %>%
  count(count_even) %>%
  ggplot(aes(x = factor(count_even), y = n, fill = n)) +  # make x categorical
  geom_bar(stat = "identity", color= "grey20") +
  scale_fill_gradient(low = "skyblue", high = "green2") +
  labs(
    title = "Even Colors Distribution",
    subtitle = "count of even colors packages",
    x = "No. Even Rows",
    y = "Packages"
  )+
  geom_text(aes(label = n), color = "black", size = 4,vjust = 0)+
  theme_minimal()
```

## Multiple Types of Samples

I will create a function that create sample for each number of colors and package size I want, and then calculate some interesting parameters

```{r function multy-sample, warning=FALSE, error=FALSE, echo=F}
mega_snack<- function(nn,n_unit,n_color)
  {
  m_sample<- length(n_unit)*length(n_color)
  nul_mat= matrix(nrow = m_sample, ncol = 6)
  res<- cbind(rep(n_unit,length(n_color)),
              sort(rep(n_color,length(n_unit))),
              nul_mat)
    colnames(res)<- c("n_unit", "n_color", "even_count", "even_evens",
                    "var_col","all_even","low_color","smallest_col")
    for (i in 1:(dim(res)[1]))
  {
    low_color<- 0.666*res[i,1]/(res[i,2])
    small_sample<- sample_MnM(nn,res[i,1],res[i,2])%>%
      as_data_frame() %>% 
      mutate(even_count= rowSums(across(everything()  , ~ .x %% 2 == 0))/res[i,2],
             #how many evens colors there are
             even_evens= (rowSums(across(c(1:n_color)  , ~ .x %% 2 == 1)) %% 2 ==0)/res[i,2],
             #are the uneven colors even
             var_col=     apply(across(c(1:n_color)), 1, var),
             #var of candy per color/ type
             smallest_col=apply(across(c(1:n_color)), 1, min),
             #lowest value in color
             all_even=    even_count >= 1,
             low_color=   rowSums(across(c(1:n_color), ~ .x <= low_color ))>=1
            )
    res[i,3:8]<- c(
      mean(small_sample$even_count),
      mean(small_sample$even_evens),
      mean(small_sample$var_col),
      mean(small_sample$all_even,  na.rm = T),
      mean(small_sample$low_color),
      min(small_sample$smallest_col))
      }

  res %>% as.data.frame()
}
```

I will make the multiple sample. parameters:
```{r Data for multiple sample, warning=FALSE, error=FALSE}
n_color<- 2:8 #Number of distinct colors in each package option
grams_op<- c(25,45,150,250,330,500,750,1000) #Weight of each package option
n_unit_op<- grams_op/gram #Total candies in each package option
nn<- 1200 #Number of packages in the sample
```

Here is some random rows:
```{r baking multiple sample, warning=FALSE, error=FALSE, echo= F}
mega_snack_1<-
  mega_snack(nn,n_unit_op,n_color) %>%
  mutate(n_unit= round(n_unit,1))

knitr::kable(mega_snack_1 %>% sample_n(5) %>% 
             mutate_at(3:7, round, 3),
             caption = "Multiple sample example rows")
```

### Deep Insight on the Data

Here are some insights:

```{r hitmap, warning=FALSE, error=FALSE, echo=F}
mega_snack_1 %>%
  ggplot(aes(x = factor((round( n_unit,1) )), y = factor(n_color ), fill = even_count)) +
  geom_tile(color= "black") +
  scale_fill_gradient(low = "orange", high = "skyblue")+
  labs(title = "Even Distinct colors Probability",
       subtitle = "avarage even count percent by number of colors and units",
       x= "Lentils",
       y= "No. of colors",
       fill= "even count") +
    geom_text(aes(label = round(even_count,2)), color = "coral4", size = 4)+
  guides(colour= "none")+
  theme_minimal()
```


```{r hitmap_2, warning=FALSE, message=F,error=FALSE, echo=F}
mega_snack_1 %>%
  ggplot(aes(x = n_unit, y = even_evens, group = n_color, color = n_color)) +
  geom_line(linewidth = 1.2) +
  scale_color_gradientn(
    colors = c("violet","purple2","deepskyblue", "green2","yellow3","orange2", "red3"),
    limits = c(2, 8),
    breaks = 2:8,
    guide = guide_colorbar(title = "colors")
  )+
  labs(
    title = "Does the Uneven Colors Even",
    subtitle = "probability of uneven distinct Lentils colors to be even, or even lentils",
    x = "Lentils",
    y = "No. of colors"
  )+
  guides(colour = guide_legend(reverse=T))+
  theme_minimal()+
  theme(
  panel.grid.major.y = element_line(color = "grey65"),
  panel.grid.minor.y = element_line(color = "grey80"),
  panel.grid.major.x = element_line(color = "grey80")   # optional, if you want to keep it clean
)
```

Here is probability of all even, and whether there is pattern.

```{r ggplot all even, warning=FALSE, message=F,error=FALSE, echo=F}
Heatmap_all_evens<-
  mega_snack_1 %>%
  ggplot(aes(x = as.factor(round(n_unit,1) ), y = n_color , color = all_even)) +
  geom_point(size = 8 ) +
  scale_color_gradient(low = "yellow", high = "darkolivegreen3")+
  labs(title = "Were all Colors Evens",
       subtitle = "probability of all distinct color of lentils been even",
       x = "Lentils",
       y = "No. of colors",
       color = "all even") +
        geom_text(aes(label = sprintf("%.1f%%", all_even * 100)), color = "black", size = 3)+
    guides(size= "none")+
  theme_bw()

ggsave(Heatmap_all_evens, 
       filename = "Heatmap_all_evens.png",
       height = 4, width = 5, units = "in")

Heatmap_all_evens
```

```{r ggplot missing color, warning=FALSE, error=FALSE, echo=F}
mega_snack_1 %>%
  ggplot(aes(x = factor(round(n_unit, 1)), y = n_color, fill = low_color))+
  geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.2f%%", low_color * 100)), color = "black", size = 4)+
  scale_fill_gradient(low = "darkolivegreen3", high = "purple3",labels = scales::percent)+
  labs(title = "Probability of Small Apearing Color",
       subtitle = "probability of at least one color apear less than 66% of expected value",
       x = "Lentils",
       y = "No. of colors",
       fill = "low color")+
  theme_minimal()
```
Here we can see the smallest % of Lentils in one color as seen in my sample:

```{r smallest color vs expected, warning=FALSE, error=FALSE, echo=F}
mega_snack_1 %>% 
  select_(1,2,8) %>% 
  mutate(smallest_to_mu= smallest_col*n_color/n_unit) %>%
   ggplot(aes(x = factor(round(n_unit, 1)), y = n_color, fill = smallest_to_mu))+
  geom_tile(color = "white") +
    geom_text(aes(label = sprintf("%.1f%%", smallest_to_mu * 100)), color = "black", size = 4)+
  scale_fill_gradient(low = "red3", high = "darkolivegreen3",labels = scales::percent)+
  labs(title = "Smallest Color as Percent of Expected Value",
       x = "Lentils",
       y = "No. of colors",
       fill = "% of Expected Value")+
  theme_bw()
```

As we can see, only the small package (less than 50 lentils) have high probability of at least one color to appear severely lower.

Therefore, splitting package by color on the big ones should be relatively even.

### Using Regression for Correlation Check

I wanted to see if there is statistic correlation of the number of distinct colors of package size to the probability of all colors have equal counts of each color. as a result, I chose to check this claim with regression. Furthermore, I did the same regression adding another potential correlated parameter: the evenness of the number of distinct colors.

```{r regression 1, warning=FALSE, message=F, error=FALSE, echo=F, cache=T}
mega_snack_open<- function(nn,n_unit,n_color)
  {
  m_sample<- nn*length(n_unit)*length(n_color)
  nul_mat= matrix(nrow = m_sample, ncol = 2)
  res<- cbind(rep(n_unit,nn*length(n_color)),
         sort(rep(n_color,nn*length(n_unit))),
              nul_mat)
  for (i in 1:(dim(res)[1]))
  {
    one_bag<- Create_bag(res[i,1],res[i,2])%>%
      as_data_frame() %>% 
      mutate(even_count= rowSums(across(everything()  , ~ .x %% 2 == 0)),
             all_even=   rowSums(across(c(1:n_color)  , ~ .x %% 2 == 0))== res[i,2],
            )
    res[i,3]<- one_bag$even_count
    res[i,4]<- one_bag$all_even
      }
  colnames(res)<- c("n_unit", "n_color", "even_count","all_even")
  res %>% as.data.frame()
  }

grams_op_2<- c(150,250,500,1000)

mega_snack_2<- mega_snack_open(60,grams_op_2,2:10)
mega_snack_2<- mega_snack_2 %>% 
  mutate(color_No2= n_color %%2 == 0)

glm_all_even<- glm(all_even ~ n_color + n_unit, 
                      data = mega_snack_2, 
                      family = binomial)
glm_all_even_2<- glm(all_even ~ n_color + n_unit + color_No2, 
                      data = mega_snack_2, 
                      family = binomial)

stargazer(glm_all_even, glm_all_even_2, type = "text")
```

As you can see, there is no statistic clear correlation of the number of colors to probability of all distinct colors to be even.

```{r regression 2, warning=FALSE, message=F, error=FALSE, echo=F, cache=T}
glm_even_count<- glm(cbind(n_color, n_color - even_count) ~ n_color + n_unit + color_No2, 
                 data = mega_snack_2, 
                 family = binomial)
glm_even_count_2<- glm(cbind(n_color, n_color - even_count) ~ n_color + n_unit, 
                 data = mega_snack_2, 
                 family = binomial)

stargazer(glm_even_count, glm_even_count_2, type = "text")
```

Like the previous regression, the second regression that examines the correlation of the number of colors and package size to the percentage of even numbers did not find a statistically significant correlation. In addition, visually there is no clear correlation.

```{r visual glm_even_count, message=F, error=FALSE, warning=F,echo=F, cache=T}
library(viridis)
library(ggridges)

mega_snack_2 %>%
    mutate(even_count_per= even_count / n_color ) %>% 
  ggplot(aes(
    x = even_count_per,
    y = factor(n_color),
    fill = stat(x)
  )) +
  geom_density_ridges_gradient(scale = 2, rel_min_height = 0.005) +
  scale_fill_viridis(name = "Even Ratio", option = "C") +
  labs(
    title = "Distribution of Even Counts by Distinct Color",
    subtitle = "Gradient color shows percent of even colors",
    x = "Even count Percent",
    y = "No. colors"
  ) +
  theme_minimal()+
  stat_summary(
    fun = mean,
    geom = "point",
    aes(group = factor(n_color)),
    color = "black",
    size = 3,
    shape = 18
  )+
      theme(
  panel.grid.major.y = element_line(color = "grey20"),
  panel.grid.major.x = element_line(color = "grey80")
    )
```

# Summary

### Data Structure

The simulation created a random samples of snack packs, which was proven to be statistically random with known $\mu$ and $\sigma^2$. I created one sample with specific size and numbers of colors using "sample_MnM", and costume multiple samples using "mega_snack". Then, I check the relevand indicators fot this project.

I found out that:

* Small packages often lack at least one color, and sometimes contain only one color.

* As the number of colors increases, the chance that all colors have even counts drops significantly.

* For medium to large packages, the probability of any one color being significantly underrepresented (less than $\frac{2}{3}$ of its expected amount) is near zero.

suggestion for any random sampler factory (like candies, Lego, toys):

1. Smaller packages need more diversity check

2. Althernatively, I would recomand calculate the amount of each type in small packages

### Main Q: Eating M&M by Two

Although there is no clear pattern to the right M&M package for all the colors to have even count,
different approach might find a clear reason for more or less couples of M&M. Here is what I did found:

The general probability of all colors to be even in 6 colored pack is 1.5% for small 50g package 2.1% for big 1000g package, and overall 1.5%, which is more than I expected.

For 5 colored pack like Skittles the average is about 2.9%

For 2 colored pack the average is 25%, so for 2 colored marshmallow bag this will be the statistics.

See all here:

```{r summary knitr, warning=FALSE, error=FALSE, echo=F}
summary_table<-
  mega_snack_1 %>%
  rename(Colors= n_color) %>% 
  group_by(Colors) %>% 
  summarise('All Even Percent'= round(mean(all_even)*100,2) ,
            .groups = 'drop') %>%
  as.data.frame()

knitr::kable(summary_table, caption = "Probability of All Colors Even by Pack Colors Number")
```

### Conclusions

To sum it up, for each medium pack the probability of all even colors is 1.4%, or 1 in a 73 packs of 250g. So I might need to change my snack preference to marshmallow if I want to keep this method.

This project allowed me to implement simulation methods in response to a real (albeit silly) question, and evaluate it statistically from end to end..

I applied:

* Simulation by demand

* Exploratory analysis

* Hypothesis testing

* Distribution checks

* Outliers detection

* Visualization using R

In addition, I created the infrastructure for similar questions with different parameters to be checked in a reusable, structured way.


